# Analysis

```{r load-data, include=FALSE}
library(scales)
library(tidyverse)
library(knitr)
project <- read_csv("Tidylearn Thesis Data - Sheet1.csv", na = c("", "?", "NA"))
```


```{r, data-manipulation, include=FALSE}
project <- project %>%
  mutate(tidyverse = ifelse(tidyverse == 0, "base R", "tidyverse"))


project$sub_analysis[is.na(project$sub_analysis)] <- 0
```

## Exploratory Data Analysis

```{r, renaming variables, echo=FALSE}
project <- project %>% 
  mutate(
    creative = create_new_var + change_var + sub_analysis + sub_data, 
    theme = eda_theme + rel_data
    )

#make rest of relevant variables to yes and no, can't figure out how to do it in
#one line
project <- project %>%
  mutate(create_new_var = ifelse(create_new_var == 0, "no", "yes")) %>%
  mutate(change_var = ifelse(change_var == 0, "no", "yes")) %>%
  mutate(sub_analysis = ifelse(sub_analysis == 0, "no", "yes")) %>%
  mutate(sub_data = ifelse(sub_data == 0, "no", "yes")) %>%
  mutate(viz_mult_mak = ifelse(viz_mult_mak == 0, "no", "yes")) %>%
  mutate(viz_mult_interpret = ifelse(viz_mult_interpret == 0, "no", "yes")) %>%
  mutate(eda_theme = ifelse(eda_theme == 0, "no", "yes")) %>%
  mutate(rel_data = ifelse(rel_data == 0, "no", "yes")) %>%
  mutate(slr_fit = ifelse(slr_fit == 0, "no", "yes")) %>%
  mutate(mlr_fit = ifelse(mlr_fit == 0, "no", "yes")) %>%
  mutate(mlr_check_cond = ifelse(mlr_check_cond == 0, "no", "yes")) %>%
  mutate(prediction = ifelse(prediction == 0, "no", "yes")) %>%
  mutate(ht = ifelse(ht == 0, "no", "yes")) %>%
  mutate(ht_check_cond = ifelse(ht_check_cond == 0, "no", "yes"))
```

As of now, the groups have an unequal amount of observations because I have gone
through more Base R projects. Just to note, the following data conclusions may 
change with time.

```{r, difference in project counts, echo=FALSE}
project %>%
  count(tidyverse) %>%
  kable()
```

Note: Currently, the creativity variable is a sum of the four different components: 
creating new variables, mutating existing ones, performing a subgroup analysis 
and subsetting the data set.

### Creativity
The creativity score is simply a sum of four different components of projects:
creating new variables, mutating existing ones, performing a subgroup analysis, 
and subsetting the data set.

In terms of total creativity, there is generally a higher proportion of more 
creativity points for Tidyverse projects.

```{r, looking at creative stat, echo=FALSE}
project %>%
  group_by(tidyverse, creative) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  kable(knitr.kable.NA = TRUE)
```

The differences in creativity scores can be elucidated through analyses into each
of the four separate variables combining to form the creativity metrics. Perhaps
the most distinct difference occurs in the creation of new variables, which is 
defined as any data manipulation throughout the EDA process where a previously 
non-existing variable is formed. As an example, one project created a new 
variable tracking if a movie had won any of the following awards: best picture,
best actor, bestactress, or best director, and that project had this variable 
coded as "yes."

```{r, looking at creating new variable, echo=FALSE}
project %>%
  group_by(tidyverse, create_new_var) %>%
  summarise(n = n()) %>%
  mutate(proportion = percent(n / sum(n))) %>%
  filter(create_new_var == "yes") %>%
  kable()
```

As shown above, it is clear that the proportion of projects with the creation of
new variables is higher in Tidyverse ones. This may be because it is very easy
to create new variables using `mutate()` in the Tidyverse.

However, one covariate in particular, mutating existing variables, did not seem 
to vary frequency-wise across different syntax users. Within the movies dataset, 
an example from some projects was to ignore G-rated movies from the dataset.

In fact, it seems as though slightly more base R studies changed pre-existing 
variables. 

```{r, looking at changing variable, echo=FALSE}
project %>%
  group_by(tidyverse, change_var) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(change_var == "yes") %>%
  kable()
```

The mechanisms for doing so are not much more complicated in base R. An example
of changing an existing studio variable to "Warner Bros. Studios" or "Other" is
listed below in both Base R and the Tidyverse for comparison:

Tidyverse:
`movies <- movies %>%
mutate(studio = (if_else(studio == "Warner Bros. Pictures", "Warner Bros. 
Pictures", "Other"))`

Base R:
`movies$studio <- if_else(movies$studio == "Warner Bros. Pictures", "Warner 
Bros. Pictures", "Other")`

The other two variables need to be re-inputted and updated, so the language and 
subsections remain below to avoid confusion.

### Subgroup Analysis

The presence of a form of a subgroup analysis was measured in regards to 
creativity. To score a 1 for this variable, projects analyzed a portion of the 
data for parts of their EDA. As an example, a project receiving a "yes" for this
category may have analyzed how the audience ratings for R rated movies compared 
to that of PG-13 movies in their bivariate analysis.

This variable has not been recoded yet.

```{r, looking at subgroup analysis variable, echo=FALSE}
project %>%
  group_by(tidyverse, sub_analysis) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(sub_analysis == "yes") %>%
  kable()

```

### Subsetting dataset

The final variable included in the creativity metric was the usage of a subset 
of the data for the entire analysis, such as analyzing solely PG movies for the 
scope of the analysis. 

This variable has also not been recoded yet fully, but it appears as though 
Tidyverse projects have a signficantly higher proportion.

```{r, looking at subsetting dataset, echo=FALSE}
project %>%
  group_by(tidyverse, sub_data) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(sub_data == "yes") %>%
  kable()

```


### Overall theme
As for a rough comparison of project quality, a new variable devised of a 
summation of a theme throughout the exploratory data analysis and the use of 
relevant and sufficient supporting data. (Could be facilitated by the 
tidyverse?)

By the use of relevant and sufficient supporting data, the covariate seeks to 
address the issue that projects may have interesting themes but lack the 
analysis and coding quality to supplement their project. Thus, a category tracking
relevant data, such as finding the correlation in a linear regression instead of
merely plotting the two variables together, was created. 

Thus far, it appears as though a higher proportion of Tidyverse 
projects reach a certain quality. This may be due to the ease in using the 
Tidyverse suit of commands in a fashion that make the project easier to code
as beginners in R.

```{r, total theme, echo=FALSE}
project %>%
  group_by(tidyverse, theme) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  kable(knitr.kable.NA = TRUE)

```

Both variables within theme reflect this trend. A present theme throughout EDA 
is present in a greater proportion of Tidyverse projects. As an example, 
analyzing the impact of movie ratings on audience scores would be a sufficient 
theme, but referencing an assortment of different predictors would not register
as one.

```{r, evaulating theme during eda, echo=FALSE}
project %>%
  group_by(tidyverse, eda_theme) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(eda_theme == "yes") %>%
  kable()
```

The significant difference in couhts could be attributed to the `select()` 
function in the Tidyverse.

Tidyverse projects also had a substantially higher margin of projects with 
relevant supplmenetary data analyses.

```{r, relevant data, echo=FALSE}
project %>%
  group_by(tidyverse, rel_data) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(rel_data == "yes") %>%
  kable()
```



### Grades NOT IMPORTANT RIGHT NOW

Not important:

When analyzing the difference in grade distributions between Tidyverse and Base 
R courses, the Base R ones are not significantly higher. Upon a closer look at 
individual semesters, though, it is clear that the Fall 2013 grade distribution 
is significantly higher than any of the others, and thus, it makes sense to 
scale the grade variable by semester. 

```{r, looking at grades, echo=FALSE}
project %>%
  group_by(tidyverse) %>%
  summarise(mean(grade, na.rm = TRUE), sd(grade, na.rm = TRUE)) %>%
  kable(knitr.kable.NA = TRUE)

ggplot(project, aes(x = tidyverse, y = grade)) +
  geom_boxplot() + 
  ggtitle("Tidyverse vs. Base R grade breakdown")

project %>%
  group_by(sem) %>% 
  summarise(mean(grade, na.rm = TRUE), sd(grade, na.rm = TRUE)) %>%
  kable(knitr.kable.NA = TRUE)

ggplot(project, aes(x = sem, y = grade)) + 
  geom_boxplot() +
  ggtitle("Grade breakdown by semeseter")
```