# Methods

Since 98 percent of the projects were submitted either as an R script or R Markdown
document, the student project code could be analyzed directly on their downloaded
submission documents. Each project was examined and coded as either a 
0, if the attribute was missing, or 1, if it was present, for 16 variables. The 
remaining three covariates identified the student projects by grade, index, and 
class.

Due to privacy concerns, since the project submissions often contained the 
members of the project group, each student project was provided with an index, 
and the names of the students in each group were removed from their submission 
document. For the dataset utilized in this thesis, projects can 
solely be identified by their assigned index. A separate dataset serves as a link 
between the individual projects and their titles.

The first few rows of the dataset compiled in this project are available below.

```{r load-data, include=FALSE}
library(scales)
library(tidyverse)
library(knitr)
library(reshape2)
library(viridis)
project <- read_csv("data/tidylearn.csv", na = c("", "?", "NA"))
```


```{r, data-manipulation, include=FALSE}
project <- project %>%
  mutate(tidyverse = ifelse(tidyverse == 0, "base R", "tidyverse"))


project$sub_analysis[is.na(project$sub_analysis)] <- 0
```

```{r, renaming variables, echo=FALSE}
project <- project %>% 
  mutate(
    creative = create_new_var + change_var + sub_analysis + sub_data, 
    theme = eda_theme + rel_data
    )
```

```{r, find summary stats for results, include = FALSE}
project %>%
  group_by(tidyverse) %>%
  summarize(
    mean_create = mean(create_new_var, na.rm = TRUE), sd_create = sd(create_new_var, na.rm = TRUE),
    mean_change = mean(change_var, na.rm = TRUE), sd_change = sd(change_var, na.rm = TRUE),
    mean_analy = mean(sub_analysis, na.rm = TRUE), sd_analy = sd(sub_analysis, na.rm = TRUE),
    mean_data = mean(sub_data, na.rm = TRUE), sd_data = sd(sub_data, na.rm = TRUE),
    mean_theme = mean(eda_theme, na.rm = TRUE), sd_theme = sd(eda_theme, na.rm = TRUE),
    mean_rel = mean(rel_data, na.rm = TRUE), sd_rel = sd(rel_data, na.rm = TRUE),
    mean_creative = mean(creative, na.rm = TRUE), sd_creative = sd(creative, na.rm = TRUE),
    mean_depth = mean(theme, na.rm = TRUE), sd_depth = sd(theme, na.rm = TRUE)
  )
```

```{r, mutate vars, echo=FALSE}
#make rest of relevant variables to yes and no, can't figure out how to do it in
#one line
project <- project %>%
  mutate(
    create_new_var = ifelse(create_new_var == 0, "no", "yes"),
    change_var = ifelse(change_var == 0, "no", "yes"),
    sub_analysis = ifelse(sub_analysis == 0, "no", "yes"),
    sub_data = ifelse(sub_data == 0, "no", "yes"),
    viz_mult_mak = ifelse(viz_mult_mak == 0, "no", "yes"),
    viz_mult_interpret = ifelse(viz_mult_interpret == 0, "no", "yes"),
    eda_theme = ifelse(eda_theme == 0, "no", "yes"),
    rel_data = ifelse(rel_data == 0, "no", "yes"),
    slr_fit = ifelse(slr_fit == 0, "no", "yes"),
    mlr_fit = ifelse(mlr_fit == 0, "no", "yes"),
    mlr_check_cond = ifelse(mlr_check_cond == 0, "no", "yes"),
    prediction = ifelse(prediction == 0, "no", "yes"),
    ht = ifelse(ht == 0, "no", "yes"),
    ht_check_cond = ifelse(ht_check_cond == 0, "no", "yes")
  )
```

```{r, display contents}
head(project)
```

The student projects were not compiled into PDF or HTML files to confirm that the code
worked, since it was a near-impossible task to determine which version of R packages
the students utilized, as some of these commands are now defunct in the most
recent versions of the packages. Because of this decision, this analysis operates
under the assumption that the code produced the desired results in each project
and did not require further debugging. 

The contents of the student project code were still analyzed for clarity, 
as well as creativeness and depth through the 16 indicator variables.

1. `index`: Project ID
1. `grade`: Score on final project
1. `sem`: Semester course taken
1. `r_rmd`: Was the submission an R script (with prose of the project turned in as a Word document) or was the submission an R Markdown file?
1. `tidyverse`: Project used `"tidyverse"` or `"base R"` syntax
1. `create_new_var`: Students created a new variable based on existing variables, `"yes"` or `"no"`,
1. `change_var`: hello
1. `sub_analysis`: hello
1. `sub_data`: hello
1. `viz_mult_mak`: hello
1. `viz_mult_interpret`: hello
1. `eda_theme`: hello
1. `rel_data`: hello
1. `slr_fit`: hello
1. `mlr_fit`: hello
1. `mlr_check_cond`: hello
1. `prediction`: hello
1. `ht`: hello
1. `ht_check_cond`: hello

## Creativity

The creativity metric seeks to encapsulate anything students coded that was 
not directly specified in the instructions but provided a purpose 
in their projects. The metric's possible scores range from 0 to 4, as a project is
scored with a single point for each of the following:

1. Creation of new variable(s) based on existing variables 
1. Transformation of existing variables 
1. Existence of a subgroup analysis
1. The use of a subset of the dataset for all steps of the project

In the case of the student group projects utilizing the Tidyverse syntax, groups
were still given scores of ones if they satisfied these conditions in Base R form.
While rare, two (NEED TO CONFIRM) groups in labs taught in the Tidyverse created
or transformed covariates using Base R syntax, which was likely due to alternative
resources, such as Stack Overflow, that at the time, did not contain many Tidyverse
solutions. Now, though, as the Tidyverse's popularity continues to grow, more online
resources incorporate and promote Tidyverse solutions.

### Creation of New Variable(s)

The creation of a new variable(s) is defined as any data manipulation throughout the
EDA process where student groups compose a previously non-existing covariate. As
an example, one project created a new variable tracking if a movie had won any of
the following awards: best picture, best actor, best actress, or best director, and 
that project had this variable coded as "yes." In order to score a 1, the student 
project also had to utilize the new variable within an aspect of their analysis. 
This condition filtered for groups that created unnecessary covariates. However, a 
score of 1 would be valid if the group did not use the variable in the inference
or regression sections, but did explore the covariate in their EDA.

### Transformation of Existing Variables

Although related to the above covariate, the transformation of existing variables
did not qualify as creating new variables, or vice versa. In this situation, a 
project would score a one if the student group mutated a variable already existing
within the dataset, generally to highlight certain cases. For instance, a few 
project groups decided to change `mpaa_rating` to either "R" or "Other," if the 
movie was not rated R. Similar to the requirements for scoring a one for the creation
of new variable(s) covariate, the mutation was required to be employed to some end,
as groups would have to provide at least a cursory analysis of the newly-mutated
variable to score a one. 

A distinction between scoring a one for this covariate and one for 
subsetting the dataset or conducting a subgroup analysis is that filtering the 
dataset for just entries that cover a portion of levels within a specific variable
would qualify as a part of either a subgroup analysis or data subset, but not this
covariate. Also, converting a factor variable that could be potentially
read in as one when loading the dataset did not qualify as a mutation of an 
existing variable for this study.


### Existence of Subgroup Analysis

The presence of a subgroup analysis was measured in regards to 
creativity. Projects that received a one analyzed portions of the data 
during their EDA process. Groups could use an assortment of commands
to satisfy a score of a 1, such as a normal boxplot, five-number summary of a
specific variable within the movies dataset, or a subsetting with a corresponding
numerical or graphical analysis. As an example, a project receiving a one for this
category may have analyzed how the audience ratings for R rated movies compared 
to that of PG-13 movies in their bivariate analysis.

### Use of a Data Subset for Project's Entirety

Although the use of a data subset covariate may seem similar to the one above,
this variable received a one for a different aspect of the final group 
projects. Here, student groups are not just using the provided movies dataset
for their EDA, inference, and regression—they are intentionally focusing on a few
characteristics of the movie dataset. Student projects were not required to employ
the same subsetted data throughout the entire analysis, but
they did have to analyze related aspects of the movies dataset to qualify for a 
one. For example, one student group scrutinized solely PG-13 rated movies for 
their final project, while another used the PG-13 rated movies subset for the EDA,
PG-13 movies released after 2000 for the inference, and then the same PG-13 rated
movies subset utilized in the EDA process for their regression analysis.

## Depth

The depth metric measures the level of depth of the analysis, both in terms of the
statistical methods utilized and in terms of story-telling. The metric ranges
from 0-2 and is scored with 1 point for each of the following: 

1. Presence of consistent theme throughout the project
2. Use of relevant data

The depth metric was created to qualify findings from the creativity score, to 
confirm that the syntax producing the more creative student projects also were of
the same (or better) quality. Although creativity is imperative in these final
projects, student groups also cannot skip parts of the data analysis cycle.

### Consistent Theme

In the world of data science, story-telling is such an important aspect, just
as story-telling is designed to be for the STA 101 final group projects. A strong 
final project requires a story: a leading question, initial findings, subsequent 
analyses, and conclusion, all formed around a specific theme. Although this 
covariate's scoring was subjective, the requirements for final projects to score
a one were similar to those defining the creativity metric. To receive a one, student
groups clearly linked the steps in their analysis, often choosing
to focus on a few aspects within the entire movie dataset. For instance,
analyzing the impact of movie ratings on audience scores would qualify as a sufficient 
theme, but referencing an assortment of different predictors would not register
as one. 

### Presence of Relevant Data

Another subjective variable, the presence of relevant data was formed to complement
the consistent theme covariate. To receive a one for this variable, student groups were
required to sufficiently use R to create insights surrounding their chosen theme(s).
The covariate addresses the issue that projects may have interesting themes but lack the 
analysis and coding quality to supplement their project. As an example, an aspect
of a group project that would have scored a one for this category would have 
displayed the correlation coefficient between two numerical variables instead of 
plotting them together and failing to acknowledge the figure in the project submission.
If the majority of the coding could be employed to support the final project, the
project group received a one.