# Results

Both numerically and graphically, the student projects dissected in this study 
using the tidyverse syntax scored higher on all three of the developed
metrics. Tidyverse projects are much more prevalent in the upper levels of 
the three variables, and their means and standard deviations are significantly 
different as well.

## Creativity Metric

```{r, difference in project counts, include=FALSE}
project %>%
  count(tidyverse) %>%
  kable()
```

Despite there being only 82 base R projects recorded to the 123 tidyverse projects,
there more base R projects scored a 0 or 1 on the creativity metric
than those using the tidyverse syntax. Overall, there was a single project that scored
a perfect four out of the 205 total projects, and the majority of the 
projects scored a one or two in the creativity metric.

However, within the Tidyverse projects, more than half (56.1 percent) registered at
least a two on creativity, compared to just 20.7 percent of base R projects.

```{r, creativity metrics stats, include=FALSE}
project %>%
  group_by(tidyverse, creative) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  kable(knitr.kable.NA = TRUE)
```


```{r, creativity metric visualization, echo=FALSE, warning = FALSE}
creative_counts <- project %>%
  group_by(tidyverse, creative) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n)))

creative_counts <- creative_counts %>%
  filter(creative == 0 | creative == 1 | creative == 2 | creative == 3 | creative == 4) %>%
  mutate(creative = as.factor(creative))

tidy_fake_data <- data.frame(tidyverse = "base R", creative = "4", n = 0, prop = "0.0%")
tidy_fake_data <- tidy_fake_data %>% mutate(tidyverse = as.character(tidyverse),
                creative = as.character(creative),
                prop = as.character(prop))

creative_counts <- data.frame(rbind(as.matrix(creative_counts),
                         as.matrix(tidy_fake_data)))

creative_counts$prop <- as.numeric(str_remove(as.character(creative_counts$prop), "%"))

ggplot(data = creative_counts, aes(x = creative, y = prop, fill = tidyverse, 
                                   group = tidyverse)) +
  geom_bar(stat = "identity", position = position_dodge(), width = .8) +
  scale_fill_viridis_d(name = "Syntax", labels = c("Base R", "Tidyverse"), begin = .37, end = .75) +
  labs(x = "Creativity Score", y = "Proportion of Projects", title = "Tidyverse Syntax Projects Score Higher on the Creativity Metric on Average")
```

The average creativity scores and corresponding standard deviations for projects
utilizing the two syntaxes are available below as well:

```{r, creativity table, echo = FALSE}
creativity_table <- data.frame(Syntax = "Base R", Mean = 1.1, SD = .6)
second <- data.frame(Syntax = "Tidyverse", Mean = 1.7, SD = .8)
creativity_table <- rbind(creativity_table, second)
kable(creativity_table)
```

| Syntax    | Mean | Standard Deviation |
|-----------|------|--------------------|
| Base R    | 1.1  | 0.6                |
| Tidyverse | 1.7  | 0.8                |

FOR MINE: Which table should I use? One has SD as Standard Deviation

The following subsections will contain a comparison of base R and tidyverse student
projects for each of the four variables combined to form the creativity metric,
as well as potential reasons for the outcomes.


### Creation of New Variable(s)

Out of the four variables that form the creativity metric, the starkes difference
between projects employing the two syntaxes was within the creation of new variable(s)
covariate. Nearly half of all final projects using the Tidyverse syntax featured
a creation of a new variable, whereas less than a quarter of base R projects did.

```{r, looking at creating new variable, echo=FALSE, warning = FALSE}
project %>%
  group_by(tidyverse, create_new_var) %>%
  summarise(n = n()) %>%
  mutate(proportion = percent(n / sum(n))) %>%
  filter(create_new_var == "yes") %>%
  kable()
```

This difference may be due to the ease in utilizing the Tidyverse's `mutate()`
function, which allows users to create new variables using a single command. In
base R, students are instructed to use `$` notation, which may be harder
to grasp due to issues with variable selection with `$` and its usage in other
base R tasks outside of just creating new variables.

### Transformation of Existing Variables

While more than 75 project groups created new variables, far less opted to mutate
existing ones in the movies dataset. Because the counts for both base R and tidyverse
projects were lower, the resulting proportions were also much smaller. Regardless,
there were more projects using the tidyverse syntax that mutated existing variables.

```{r, looking at changing variable, echo=FALSE}
project %>%
  group_by(tidyverse, change_var) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(change_var == "yes") %>%
  kable()
```

The difference in proportions was not nearly as significant as it was for the creation
of new variables and only differed marginally. This limited difference
can be attributed to two factors. First, with the option to create their own variables,
groups may not find transforming existing variables as appealing. Second, the
mechanism to do so does not differ dramatically between the two syntaxes. For comparison,
an example of changing an existing studio variable to "Warner Bros. Studios" or
"Other" is listed in both Base R and the Tidyverse:

Base R:
`movies$studio <- if_else(movies$studio == "Warner Bros. Pictures", "Warner 
Bros. Pictures", "Other")`

Tidyverse:
`movies <- movies %>%
mutate(studio = (if_else(studio == "Warner Bros. Pictures", "Warner Bros. 
Pictures", "Other"))`

### Existence of Subgroup Analysis

Out of all the covariates forming the creativity metric, the subgroup analysis
was the most popular one satisfied for both base R and tidyverse projects. 93.5 
percent of tidyverse projects performed a type of subgroup analysis, while more 
than 3/4 of base R projects did.

```{r, looking at subgroup analysis variable, echo=FALSE, warning = FALSE}
project %>%
  group_by(tidyverse, sub_analysis) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(sub_analysis == "yes") %>%
  kable()
```

The popularity of this aspect within group projects may be explained by the copious
ways student groups could perform a subgroup analysis. However,
subgroup analyses may be easier to perform in the tidyverse due to the presence 
of the `group_by()` command, which is usually one of the most popular functions
for beginning R users in the tidyverse. In contrast, base R's `by()` function
is not nearly as well-known.

### Use of a Data Subset for Project's Entirety

The use of a subset of the provided movies dataset for the entire final project was
also significantly more popular amongst tidyverse projects, as more than 15 percent
of student projects using the tidyverse satisfied a score of a one for this covariate,
compared to less than 5 percent of all base R projects.

```{r, looking at subsetting dataset, echo=FALSE, warning = FALSE}
project %>%
  group_by(tidyverse, sub_data) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(sub_data == "yes") %>%
  kable()
```

Since there is little difference between the tidyverse's `filter()` and base R's
`subset()` functions, this discrepancy might not be able to be explained by in
an obvious manner. Perhaps, though, student groups using the tidyverse may have 
been more encouraged to utilize data subsets throughout their projects more often
than those in base R because other aspects of EDA have been performed more frequently
in tidyverse projects, so those groups may have gained additional insights that 
led them to use a data subset that base R project groups did not discover.

## Depth Metric

Although the depth metrics scores between final projects that employed base R 
syntax compared to those using the tidyverse do not provide a similar discrepancy
as the creativity score, there is still a considerable difference in the depth metric
distributions between projects of the two syntaxes.

48.8 percent of the projects using the tidyverse syntax scored a perfect two in the 
depth metric compared to 34.1 percent of base R projects, while the extreme majority
of base R projects scored a 0 or 1 in the depth metric. Overall, most projects scored
a one.

```{r, depth stats, include=FALSE}
project %>%
  group_by(tidyverse, theme) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  kable(knitr.kable.NA = TRUE)
```

```{r, depth visualization, echo=FALSE}
project_counts <- project %>%
  group_by(tidyverse, theme) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n)))

project_counts <- project_counts %>% 
  filter(theme == 0 | theme == 1 | theme == 2) %>%
  mutate(theme = as.factor(theme))

project_counts$prop <- as.numeric(str_remove(as.character(project_counts$prop), "%"))

ggplot(data = project_counts, aes(x = theme, y = prop, fill = tidyverse, group = tidyverse)) +
  geom_bar(stat = "identity", position = "dodge", width = .8) +
  scale_fill_viridis_d(name = "Syntax", labels = c("Base R", "Tidyverse"), begin = .37, end = .75) +
  labs(x = "Depth Metric", y = "Proportion of Projects", title = "Tidyverse Syntax Projects Score Higher on the Depth Metric on Average")
```

The average depth scores and corresponding standard deviations for projects
utilizing the two syntaxes are available below as well:

```{r, depth table, echo = FALSE}
depth_table <- data.frame(Syntax = "Base R", Mean = 1.2, SD = .7)
second <- data.frame(Syntax = "Tidyverse", Mean = 1.4, SD = .7)
depth_table <- rbind(depth_table, second)
kable(depth_table)
```

FOR MINE: Which table should I use? One has SD as Standard Deviation

| Syntax    | Mean | Standard Deviation |
|-----------|------|--------------------|
| Base R    | 1.2  | 0.7                |
| Tidyverse | 1.4  | 0.7                |

### Consistent Theme

Within the theme metric, the variable tracking the presence of a consistent theme
showcased a larger difference in proportions between base R and tidyverse student
projects. Compared to base R projects, where 57.3 percent boast a consistent theme,
71.5 percent of all tidyverse projects maintain consistency theme-wise.

```{r, evaulating theme during eda, echo=FALSE, warning = FALSE}
project %>%
  group_by(tidyverse, eda_theme) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(eda_theme == "yes") %>%
  kable()
```

The difference in proportions may potentially be attributed to the prevalence of
`select()` and `filter()` within the tidyverse syntax, which allows users to 
choose specific columns within a data that satisfy a certain criteria, compared 
to base R, when student groups commonly utilized square brackets to provide those
same criteria.

### Presence of Relevant Data

Proportions for the presence of relevant data covariate were very similar, with 
just a 4.4 percent difference in proportions (64.2 for Tidyverse compared to 59.8
for base R projects) between projects using the two syntaxes.

```{r, relevant data, echo=FALSE}
project %>%
  group_by(tidyverse, rel_data) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(rel_data == "yes") %>%
  kable()
```

The presence of relevant supporting data should not be greatly impacted by the 
particular coding syntax student groups employed, which may explain the small
difference in proportions for this variable.

## Multivariate Visualization Metric

Between projects using the two different syntaxes, the most significant difference in
any of the three metrics occurred within the multivariate visualization metric, where
just one final project in base R even displayed a plot containing at least three
different variables. The majority of all projects did not include multivariate
visualizations, but projects using the tidyverse syntax were more popular within
higher scores of the metric in general, with 13 percent scoring a perfect two, 
while no projects in base R reached the same score.

```{r, multi viz stats, include=FALSE, warning = FALSE}
project %>%
  group_by(tidyverse, multiviz) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  kable(knitr.kable.NA = TRUE)
```

```{r, multi viz visualization, echo=FALSE, warning = FALSE}
viz_counts <- project %>%
  group_by(tidyverse, multiviz) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n)))

viz_counts <- viz_counts %>%
  filter(multiviz == 0 | multiviz == 1 | multiviz == 2) %>%
  mutate(multiviz = as.factor(multiviz))

tidy_fake_data <- data.frame(tidyverse = "base R", multiviz = "2", n = 0, prop = "0.0%")

viz_counts <- data.frame(rbind(as.matrix(viz_counts),
                         as.matrix(tidy_fake_data)))

viz_counts$prop <- as.numeric(str_remove(as.character(viz_counts$prop), "%"))

ggplot(data = viz_counts, aes(x = multiviz, y = prop, fill = tidyverse, 
                                   group = tidyverse)) +
  geom_bar(stat = "identity", position = position_dodge(), width = .8) +
  scale_fill_viridis_d(name = "Syntax", labels = c("Base R", "Tidyverse"), begin = .37, end = .75) +
  labs(x = "Multivariate Visualization Metric", y = "Proportion of Projects", title = "Tidyverse Syntax Projects Score Higher on the Multivariate Visualization Metric on Average")
```

The average multivariate visualization scores and corresponding standard 
deviations for projects utilizing the two syntaxes are available below as well. 
Although the statistics do not represent the difference in the metric between projects
of the two syntaxes, there is still a distinct divide.

```{r, multiviz table, echo = FALSE}
multiviz_table <- data.frame(Syntax = "Base R", Mean = 0.0, SD = .1)
second <- data.frame(Syntax = "Tidyverse", Mean = 0.4, SD = .7)
multiviz_table <- rbind(multiviz_table, second)
kable(multiviz_table)
```

FOR MINE: Which table should I use? One has SD as Standard Deviation

| Syntax    | Mean | Standard Deviation |
|-----------|------|--------------------|
| Base R    | 0.0  | 0.1                |
| Tidyverse | 0.4  | 0.7                |

### Presence of a Visualization with 3+ Variables

Only one project coded in solely base R displayed a visualization with at least
three variables, compared to 30 tidyverse projects. Percentage-wise, too,
the tidyverse projects were more likely to contain a multivariate
visualization.

```{r, multiviz presence, echo=FALSE}
project %>%
  group_by(tidyverse, viz_mult_mak) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(viz_mult_mak == "yes") %>%
  kable()
```

These results may be due to the differences in base R's `plot()` function
relative to the tidyverse's `ggplot()`. Whereas `plot()` requires extra commands
to add graphical aesthetics besides x- and y-variables, the `aes()` command embedded
within `ggplot()` provides an easy platform for tidyverse users to employ other 
aesthetics besides just the x- and y-variables. For example, code is listed below
that would satisfy a score of one when using the two syntaxes.

Base R:

`movies.color <- rep("pink", length(movies$best_pic_nom))
movies.color[movies$best_pic_nom == "yes"] <- "blue"
plot(audience_score ~ critics_score, col = movies.color, data = movies)
legend(col = c("pink", "blue"))`

Tidyverse:

`ggplot(data = movies, aes(x = critics_score, y = audience_score, color = best_pic_nom)) +
    geom_point()`

FOR MINE: This code is correct, right?

### Interpretation of Multivariate Visualization

No student project using solely base R interpreted a multivariate plot, whereas
13 percent of projects utilizing the tidyverse did. Although the proportion satisfying
a score of a one is not very high, more than half of tidyverse projects that formed
a visualization with at least three variables (16 out of 30) contained sufficient
intepretations of the outputs.

```{r, multiviz interpret, echo=FALSE, warning = FALSE}
project %>%
  group_by(tidyverse, viz_mult_interpret) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(viz_mult_interpret == "yes") %>%
  kable()
```

Theoretically, the difference in proportions between projects primarily using base R 
compared to the tidyverse should not drastically differ. However, since there were
more tidyverse projects that contained multivariate plots, there was bound to
be a difference in this covariate. In general, though, the plots should not deviate
for interpretation, though some users may find comfort in viewing graphs provided
by their most commonly-used syntax.