# Results

Both numerically and graphically, the student projects using the tidyverse syntax
in this setting score higher on the creativity and depth metrics. Tidyverse projects
overwhelmingly score higher in the upper levels of the two variables, and their
mean and standard deviations are significantly different:

Creativity:

Syntax    | Mean  | Standard Deviation
--------- | ----- -------------
Base R    | 1.1   | .6
Tidyverse | 1.7   | .8

Depth:

Syntax    | Mean  | Standard Deviation
--------- | ----- -------------
Base R    | 1.2   | .7
Tidyverse | 1.4   | .7

## Creativity Metric

```{r, difference in project counts, include=FALSE}
project %>%
  count(tidyverse) %>%
  kable()
```

Despite there being only 82 base R projects recorded to the 123 Tidyverse projects,
there still were more base R projects scoring a 0 or 1 on the creativity metric
than those within the Tidyverse. Overall, there was a single project that scored
a perfect four, and the majority of the total projects scored a one or two.

Yet within the Tidyverse projects, more than half (56.1 percent) registered at
least a two on creativity, compared to just 20.7 percent of base R final projects.

The complete distribution is listed below:

```{r, looking at creative stat, echo=FALSE}
project %>%
  group_by(tidyverse, creative) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  kable(knitr.kable.NA = TRUE)


creative_counts <- project %>%
  group_by(tidyverse, creative) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n)))

creative_counts <- creative_counts %>%
  filter(creative == 0 | creative == 1 | creative == 2 | creative == 3 | creative == 4) %>%
  mutate(creative = as.factor(creative))

tidy_fake_data <- data.frame(tidyverse = "tidyverse", creative = "4", n = 0, prop = "0.0%")
tidy_fake_data <- tidy_fake_data %>% mutate(tidyverse = as.character(tidyverse),
                creative = as.character(creative),
                prop = as.character(prop))

creative_counts <- data.frame(rbind(as.matrix(creative_counts),
                         as.matrix(tidy_fake_data)))

creative_counts$prop <- as.numeric(str_remove(as.character(creative_counts$prop), "%"))

ggplot(data = creative_counts, aes(x = creative, y = prop, fill = tidyverse, 
                                   group = tidyverse)) +
  geom_bar(stat = "identity", position = position_dodge(), width = .8) +
  scale_fill_viridis_d(name = "Syntax", labels = c("Base R", "Tidyverse"), begin = .37, end = .75) +
  labs(x = "Creativity Score", y = "Proportion of Projects", title = "Tidyverse Syntax Projects Score Higher on the Creativity Metric on Average")

ggsave("Creativity.png")
```

The following subsections will dissect the similarities and differences between
base R and tidyverse student projects for the four variables combined to form the
creativity metric, as well as postulate reasons for the distributions.


### Creation of New Variable(s)

The largest difference in proportions for the four variables that make up the creativity
metric is within the creation of new variable category. Nearly half of all final 
projects using the Tidyverse syntax featured a creation of a new variable, whereas
less than a quarter of base R projects did.

```{r, looking at creating new variable, echo=FALSE}
project %>%
  group_by(tidyverse, create_new_var) %>%
  summarise(n = n()) %>%
  mutate(proportion = percent(n / sum(n))) %>%
  filter(create_new_var == "yes") %>%
  kable()
```

This difference may be due to the easability in utilizing the Tidyverse's `mutate()`
function, which allows users to create new variables using just one command. In
Base R, students are instructed to use `$` notation, which may be construed as harder
to grasp due to issues with variable selection with `$` and its usage in other
base R tasks than just creating new variables.


In fact, it seems as though slightly more base R studies changed pre-existing 
variables. 

```{r, looking at changing variable, echo=FALSE}
project %>%
  group_by(tidyverse, change_var) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(change_var == "yes") %>%
  kable()
```

The mechanisms for doing so are not much more complicated in base R. An example
of changing an existing studio variable to "Warner Bros. Studios" or "Other" is
listed below in both Base R and the Tidyverse for comparison:

Tidyverse:
`movies <- movies %>%
mutate(studio = (if_else(studio == "Warner Bros. Pictures", "Warner Bros. 
Pictures", "Other"))`

Base R:
`movies$studio <- if_else(movies$studio == "Warner Bros. Pictures", "Warner 
Bros. Pictures", "Other")`

The other two variables need to be re-inputted and updated, so the language and 
subsections remain below to avoid confusion.

### Subgroup Analysis

The presence of a form of a subgroup analysis was measured in regards to 
creativity. To score a 1 for this variable, projects analyzed a portion of the 
data for parts of their EDA. As an example, a project receiving a "yes" for this
category may have analyzed how the audience ratings for R rated movies compared 
to that of PG-13 movies in their bivariate analysis.

This variable has not been recoded yet.

```{r, looking at subgroup analysis variable, echo=FALSE}
project %>%
  group_by(tidyverse, sub_analysis) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(sub_analysis == "yes") %>%
  kable()

```

### Subsetting dataset

The final variable included in the creativity metric was the usage of a subset 
of the data for the entire analysis, such as analyzing solely PG movies for the 
scope of the analysis. 

This variable has also not been recoded yet fully, but it appears as though 
Tidyverse projects have a signficantly higher proportion.

```{r, looking at subsetting dataset, echo=FALSE}
project %>%
  group_by(tidyverse, sub_data) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(sub_data == "yes") %>%
  kable()

project %>% 
  group_by(tidyverse) %>%
  summarise(mean = mean(sub_data), sd = sd(sub_data))
```


### Overall theme
As for a rough comparison of project quality, a new variable devised of a 
summation of a theme throughout the exploratory data analysis and the use of 
relevant and sufficient supporting data. (Could be facilitated by the 
tidyverse?)

By the use of relevant and sufficient supporting data, the covariate seeks to 
address the issue that projects may have interesting themes but lack the 
analysis and coding quality to supplement their project. Thus, a category tracking
relevant data, such as finding the correlation in a linear regression instead of
merely plotting the two variables together, was created. 

Thus far, it appears as though a higher proportion of Tidyverse 
projects reach a certain quality. This may be due to the ease in using the 
Tidyverse suit of commands in a fashion that make the project easier to code
as beginners in R.

```{r, total theme, echo=FALSE}
project %>%
  group_by(tidyverse, theme) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  kable(knitr.kable.NA = TRUE)

project_counts <- project %>%
  group_by(tidyverse, theme) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n)))

project_counts <- project_counts %>% 
  filter(theme == 0 | theme == 1 | theme == 2) %>%
  mutate(theme = as.factor(theme))

project_counts$prop <- as.numeric(str_remove(as.character(project_counts$prop), "%"))

ggplot(data = project_counts, aes(x = theme, y = prop, fill = tidyverse, group = tidyverse)) +
  geom_bar(stat = "identity", position = "dodge", width = .8) +
  scale_fill_viridis_d(name = "Syntax", labels = c("Base R", "Tidyverse"), begin = .37, end = .75) +
  labs(x = "Depth Metric", y = "Proportion of Projects", title = "Tidyverse Syntax Projects Score Higher on the Depth Metric on Average")
ggsave("Depth.png")
```

HERE'S THE DEPTH TABLE:

Syntax    | Mean  | Standard Deviation
--------- | ----- -------------
Base R    | 1.2   | .7
Tidyverse | 1.4   | .8


Both variables within theme reflect this trend. A present theme throughout EDA 
is present in a greater proportion of Tidyverse projects. As an example, 
analyzing the impact of movie ratings on audience scores would be a sufficient 
theme, but referencing an assortment of different predictors would not register
as one.

```{r, evaulating theme during eda, echo=FALSE}
project %>%
  group_by(tidyverse, eda_theme) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(eda_theme == "yes") %>%
  kable()
```

The significant difference in couhts could be attributed to the `select()` 
function in the Tidyverse.

Tidyverse projects also had a substantially higher margin of projects with 
relevant supplmenetary data analyses.

```{r, relevant data, echo=FALSE}
project %>%
  group_by(tidyverse, rel_data) %>%
  summarise(n = n()) %>%
  mutate(prop = percent(n / sum(n))) %>%
  filter(rel_data == "yes") %>%
  kable()
```



### Grades NOT IMPORTANT RIGHT NOW

Not important:

When analyzing the difference in grade distributions between Tidyverse and Base 
R courses, the Base R ones are not significantly higher. Upon a closer look at 
individual semesters, though, it is clear that the Fall 2013 grade distribution 
is significantly higher than any of the others, and thus, it makes sense to 
scale the grade variable by semester. 

```{r, looking at grades, echo=FALSE}
project %>%
  group_by(tidyverse) %>%
  summarise(mean(grade, na.rm = TRUE), sd(grade, na.rm = TRUE)) %>%
  kable(knitr.kable.NA = TRUE)

ggplot(project, aes(x = tidyverse, y = grade)) +
  geom_boxplot() + 
  ggtitle("Tidyverse vs. Base R grade breakdown")

project %>%
  group_by(sem) %>% 
  summarise(mean(grade, na.rm = TRUE), sd(grade, na.rm = TRUE)) %>%
  kable(knitr.kable.NA = TRUE)

ggplot(project, aes(x = sem, y = grade, color = tidyverse)) + 
  geom_boxplot() +
  ggtitle("Grade breakdown by semeseter")
```