# Educational Resources

After analyzing the retrospective study, we provided two resources intended for use
in
statistical instruction to reflect our insights in favor of the tidyverse syntax.
We altered current STA 101 labs to further adhere to the GAISE manual and provided
two code samples with explanations for `infer`, an R package created to perform 
inference tasks using a tidy framework.

## Infer Vignettes 

Currently, the tidyverse contains packages to clean, mutate, model and visualize
data, but not to perform inference tasks. 
Due to the popularity and demand for the tidyverse framework, a group of professors
created a relatively new package titled `infer`, which is fashioned as the 
tidyverse solution to inference. Like packages such as `dplyr` and `ggplot2`, 
`infer` relies on a few commands to execute a variety of tasks and benefits from
the piping structure used throughout the tidyverse. 

Unfortunately, `infer` does not presently include detailed examples of the package's
usage. Therefore, we decided to write two vignettes—longer code 
samples embedded within data analysis stories—to provide `infer` users with vivid
examples and explanations for proper usage of the package.

By inference, the two primary endeavors are performing hypothesis tests and creating
confidence intervals—hence a vignette for each. The vignettes also utilize separate
aspects of the same dataset, to provide users the comfort of a consistent data 
source while exposing them to tasks using both numerical and categorical variables.

The complete vignettes are available in the Appendix.

## Lab Enhancements

Based on the results gleaned from analyzing the student project data from 
introductory statistics courses at Duke University, we decided to fully apply 
tidyverse syntax and methods to introductory statistics labs. OpenIntro Statistics
is an open-source textbook approved for use at the undergraduate level by the
American Institute of Mathematics, and the labs accompanying OpenIntro textbooks
are available online and are used by students at many institutions, including at
Duke University. These labs were updated in 2016 to incorporate tidyverse 
syntax for data visualization and wrangling, though statistical inference still 
relied on Base R syntax. As part of this project, we have updated the code 
introduced in the labs to fully leverage the Tidyverse ecosystem, including the 
`infer` package.

Since the most recent update of the labs to incorporate tidyverse syntax, 
recommendations around introducing data visualization with `ggplot2` have changed.
Previous practice used the `qplot()` (quick plot) function, which has a simpler 
API than the `ggplot()` function, but is more cumbersome to produce complex 
multivariate visualizations with. With this update, we have completely abandoned
the use of `qplot()` and replaced it with `ggplot()`, resulting in changes in 
associated code. Two main reasons for this change are the existence
of plethora of resources for debugging `ggplot()` as well as ease of expansion to
complex visualizations.

Another major update was in the labs that focus on statistical inference. Previous
versions of the labs used a custom function called `inference()` from the `oilabs`
package, the R package used to supplement OpenIntro. This function, designed with
the best intentions in mind for highlighting the unified nature of statistical 
inference across various hypothesis tests and confidence intervals introduced in 
introductory statistics curricula, over time morphed into a function that is too 
extensive for efficient debugging and too customized for use beyond the introductory
statistics classroom. The infer package, released in 2018, was heavily inspired 
by the `inference()` function, but significantly improved the API for tidy 
statistical inference and tied it closely to how both we and the GAISE recommend
introducing statistical inference in introductory statistics curricula. The goal
of this package “is to perform inference using an expressive statistical grammar
that coheres with the tidy design framework," as per its R documentation.

For instance, here is code for generating a two-sided hypothesis test using the
two methods, with the `inference()` function first.

```{r, hypothesis test inference, eval=FALSE}
inference(y = y_variable, x = x_variable, data = data, statistic = "mean", type = "ht", null = 0, alternative = "twosided", method = “theoretical")
```

```{r, hypothesis test, eval=FALSE}
obs_diff <- data %>%
  specify(dependent ~ independent) %>% 
  calculate(stat = "diff in means", order = c("yes", "no"))
null <- data %>% 
  specify(dependent ~ independent) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
null %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")
```

And here is an example of code creating a 95 percent confidence interval using 
`inference()` and `infer`, respectively.

```{r, ci inference, eval=FALSE}
inference(y = response, data = data, statistic = "proportion", type = "ci", method = "theoretical", success = "yes")
```

```{r, ci, eval=FALSE}
data %>%
  specify(formula = text_ind ~ NULL, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = .95)
```

The labs focusing on statistical inference have thus been updated with the current
infer syntax in order to provide students with a workflow they can extend outside
of their classroom setting alongside the tidyverse, as well as one with available
online resources.

A few of the labs also featured datasets that have either been out of use for the 
past few years and will continue to be considered outmoded, were convoluted and 
required supplementary information, or intentionally or unintentionally promoted
gender stereotypes. These datasets, and the corresponding labs, were updated with
more robust and interesting material, as summarized in the table below. The 
datasets were analyzed prior to insertion to affirm that they would have the 
bandwidth to seamlessly fit the particular focus of the labs.

Previous Dataset  |  Reason for Change |  New Dataset
--------------------------- | --------------------------- |--------------------------- |
Body Dimensions | Focuses on weight comparison across binary genders  | Fast Food Nutritional Facts 
Atheism and Religion | Unverified Dataset | Youth Risk Behavior Surveillance System
Baseball | Required further dataset explanation | Human Freedom Index
North Carolina Births | Dated + Not relevant for most undergrads | Youth Risk Behavior Surveillance System


